Draft Final Report: Measles risk assessment, modelling, and benefit--cost analyses for New Zealand

### 2015-04-24 draftfinalreport_v3.tex
with lots maps (THIS IS THE ONE WE COULD KILL IF WE WANT THE MAIN MAPS GONE)

\captionsetup{width=1\textwidth}

@@

% \begin{table}[hbtp]\footnotesize
% <<results=tex,echo=FALSE>>=
% colnames(propvCohort) <- c("Age range", "Proportion")
% latex(propvCohort, file="", table.env=FALSE,rowname=NULL)
% @
% \caption{Proportions of the na\"{\i}ve population the would be covered if vaccinated by age groups}
% \label{table:vaccine_cohort}
% \end{table}

@@

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-3 package
% Tue Apr 21 10:02:47 2015
\begin{tabular}{lrrrrrr}
  \hline
 & Df & Deviance & Resid. Df & Resid. Dev & F & Pr($>$F) \\ 
  \hline
NULL &  &  & 59 & 1532.80 &  &  \\ 
  Age & 4 & 1345.17 & 55 & 187.63 & 204.01 & 0.0000 \\ 
  Ethnicity & 3 & 19.88 & 52 & 167.75 & 4.02 & 0.0141 \\ 
  NZDep & 2 & 17.10 & 50 & 150.65 & 5.19 & 0.0102 \\ 
  Age:Ethnicity & 12 & 80.80 & 38 & 69.85 & 4.09 & 0.0004 \\ 
   \hline
\end{tabular}\caption{Significance of different predictor variables for measles risk factors from 2007-2014, estimated using \autoref{eq:reg} and a quasipoisson error structure}
\label{table:anova}
\end{table}

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-3 package
% Tue Apr 21 10:02:47 2015
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -6.1644 & 0.1267 & -48.65 & 0.0000 \\ 
  Age2-4 & -1.0351 & 0.1972 & -5.25 & 0.0000 \\ 
  Age5-17 & -0.9954 & 0.1368 & -7.28 & 0.0000 \\ 
  Age18-24 & -1.7204 & 0.1926 & -8.93 & 0.0000 \\ 
  Age25+ & -3.1697 & 0.1622 & -19.55 & 0.0000 \\ 
  EthnicityAsian & 0.2464 & 0.3168 & 0.78 & 0.4415 \\ 
  EthnicityMaori & 0.2081 & 0.2121 & 0.98 & 0.3327 \\ 
  EthnicityPacific & 1.2200 & 0.2134 & 5.72 & 0.0000 \\ 
  NZDep4-7 & -0.2609 & 0.0948 & -2.75 & 0.0090 \\ 
  NZDep8-10 & -0.3043 & 0.1040 & -2.93 & 0.0058 \\ 
  Age2-4:EthnicityAsian & -2.3148 & 1.3319 & -1.74 & 0.0903 \\ 
  Age5-17:EthnicityAsian & -1.2453 & 0.4566 & -2.73 & 0.0096 \\ 
  Age18-24:EthnicityAsian & -1.0187 & 0.5743 & -1.77 & 0.0841 \\ 
  Age25+:EthnicityAsian & 0.2344 & 0.4298 & 0.55 & 0.5887 \\ 
  Age2-4:EthnicityMaori & -0.1013 & 0.3644 & -0.28 & 0.7826 \\ 
  Age5-17:EthnicityMaori & -0.1032 & 0.2521 & -0.41 & 0.6846 \\ 
  Age18-24:EthnicityMaori & -0.5723 & 0.4409 & -1.30 & 0.2022 \\ 
  Age25+:EthnicityMaori & -1.0349 & 0.5127 & -2.02 & 0.0506 \\ 
  Age2-4:EthnicityPacific & -0.9798 & 0.4676 & -2.10 & 0.0429 \\ 
  Age5-17:EthnicityPacific & -1.5709 & 0.3312 & -4.74 & 0.0000 \\ 
  Age18-24:EthnicityPacific & -1.0354 & 0.5003 & -2.07 & 0.0453 \\ 
  Age25+:EthnicityPacific & -0.6629 & 0.4279 & -1.55 & 0.1296 \\ 
   \hline
\end{tabular}\caption{Summary of the regression model results using \autoref{eq:reg} and a quasipoisson error structure, for measles risk factors for cases from 2007-12 June 2014}
\label{table:regression_riskfactors}
\end{table}

@@

@@

The regression analyses suggest that age is a particularly strong risk factor for measles. This comes as no surpise to epidemiologists or health care providers. However, our analyses also highlight other groups that are at greater risk of measles. In particular, Pacific people are at greater risk per capita, as are the less deprived (NZDep 1--3), and European and Maori 5--17 year old children compared to Asian and Pacific ethnicity children of the same age. Furthermore, Pacific origin people of all categories from 2--24 years old were less likely on a per capita basis to be cases than Europeans or Maori. Interpretation of these results must still be viewed with some caution, however, as there is very likely a spatial effect that might not be accounted for in these analyses. For example, increased numbers of more wealthy Europeans were measles cases, as was true for Maori. However, the alternative situation was clear for Pacific people, especially of the 0--1 year olds (\autoref{fig:percap_all} and \autoref{fig:cases_all}). Results such as this may simply be due to the spatial effects in the communities the outbreaks were occurring in, rather than an effect of wealth. The risk maps using the spatially explicit NIR vaccination data suggest that there is some heterogeneity in the vaccination cover at the area unit (a level above meshblock) for both MMR1 and MMR2 (see histograms within \autoref{fig:fig12006}, \autoref{fig:fig12007}, \autoref{fig:fig12008}, \autoref{fig:fig12009}, \autoref{fig:fig12010}, \autoref{fig:fig12011}, \autoref{fig:fig12012}, \autoref{fig:fig12013}, \autoref{fig:fig12014}, \autoref{fig:fig22006}, \autoref{fig:fig22007}, \autoref{fig:fig22008}, \autoref{fig:fig22009}, \autoref{fig:fig22010}, \autoref{fig:fig22011}, \autoref{fig:fig22012}, \autoref{fig:fig22013}, and \autoref{fig:fig22014}).

@@

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.9\textwidth]{percap_eth_age.pdf}
     \end{center}
     \caption{Per capita measles case rates by prioritized ethnicity and age from 2007--2014. Full details are given in (\autoref{fig:percap_all})}
     \label{fig:PerCapAgeEth}
\end{figure}

>>>>

\captionsetup{width=1\textwidth, font={tiny}}

@@

@@

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-3 package
% Fri Apr 24 16:29:15 2015
\begin{tabular}{lrrrrrr}
  \hline
 & Df & Deviance & Resid. Df & Resid. Dev & F & Pr($>$F) \\ 
  \hline
NULL &  &  & 59 & 1532.80 &  &  \\ 
  Age & 4 & 1345.17 & 55 & 187.63 & 204.01 & 0.0000 \\ 
  Ethnicity & 3 & 19.88 & 52 & 167.75 & 4.02 & 0.0141 \\ 
  NZDep & 2 & 17.10 & 50 & 150.65 & 5.19 & 0.0102 \\ 
  Age:Ethnicity & 12 & 80.80 & 38 & 69.85 & 4.09 & 0.0004 \\ 
   \hline
\end{tabular}\caption{Significance of different predictor variables for measles risk factors from 2007-2014, estimated using \autoref{eq:reg} and a quasipoisson error structure}
\label{table:anova}
\end{table}

\vspace{5mm} %5mm vertical space
\begin{table}
% latex table generated in R 3.0.2 by xtable 1.7-3 package
% Fri Apr 24 16:29:15 2015
\begin{tabular}{rrrrr}
  \hline
 & Estimate & Std. Error & t value & Pr($>$$|$t$|$) \\ 
  \hline
(Intercept) & -6.1644 & 0.1267 & -48.65 & 0.0000 \\ 
  Age2-4 & -1.0351 & 0.1972 & -5.25 & 0.0000 \\ 
  Age5-17 & -0.9954 & 0.1368 & -7.28 & 0.0000 \\ 
  Age18-24 & -1.7204 & 0.1926 & -8.93 & 0.0000 \\ 
  Age25+ & -3.1697 & 0.1622 & -19.55 & 0.0000 \\ 
  EthnicityAsian & 0.2464 & 0.3168 & 0.78 & 0.4415 \\ 
  EthnicityMaori & 0.2081 & 0.2121 & 0.98 & 0.3327 \\ 
  EthnicityPacific & 1.2200 & 0.2134 & 5.72 & 0.0000 \\ 
  NZDep4-7 & -0.2609 & 0.0948 & -2.75 & 0.0090 \\ 
  NZDep8-10 & -0.3043 & 0.1040 & -2.93 & 0.0058 \\ 
  Age2-4:EthnicityAsian & -2.3148 & 1.3319 & -1.74 & 0.0903 \\ 
  Age5-17:EthnicityAsian & -1.2453 & 0.4566 & -2.73 & 0.0096 \\ 
  Age18-24:EthnicityAsian & -1.0187 & 0.5743 & -1.77 & 0.0841 \\ 
  Age25+:EthnicityAsian & 0.2344 & 0.4298 & 0.55 & 0.5887 \\ 
  Age2-4:EthnicityMaori & -0.1013 & 0.3644 & -0.28 & 0.7826 \\ 
  Age5-17:EthnicityMaori & -0.1032 & 0.2521 & -0.41 & 0.6846 \\ 
  Age18-24:EthnicityMaori & -0.5723 & 0.4409 & -1.30 & 0.2022 \\ 
  Age25+:EthnicityMaori & -1.0349 & 0.5127 & -2.02 & 0.0506 \\ 
  Age2-4:EthnicityPacific & -0.9798 & 0.4676 & -2.10 & 0.0429 \\ 
  Age5-17:EthnicityPacific & -1.5709 & 0.3312 & -4.74 & 0.0000 \\ 
  Age18-24:EthnicityPacific & -1.0354 & 0.5003 & -2.07 & 0.0453 \\ 
  Age25+:EthnicityPacific & -0.6629 & 0.4279 & -1.55 & 0.1296 \\ 
   \hline
\end{tabular}\caption{Summary of the regression model results using \autoref{eq:reg} and a quasipoisson error structure, for measles risk factors for cases from 2007-12 June 2014}
\label{table:regression_riskfactors}
\end{table}

@@

Vaccination coverage (0--13 years) and serosurvey estimates of immunity ($>13$ years) among different age classes provided by the Ministry were also used to map the immunity among the 2004 2013 area units for those ages not included on the NIR database. Vaccine efficacy for 0--4 year olds was not necessary for these analyses, for those $\leq$ 13 years 99\% vaccine efficacy was used (\autoref{table:sero}) and for $\geq$ 14 year olds equivocal serological results were considered non-immune (\autoref{table:sero}). Again there was heterogeneity at the local level (\autoref{fig:sero_imm_map}) with pockets of the nation and cities with lower coverage than other areas (\autoref{fig:sero_imm_a}, \autoref{fig:sero_imm_w}, and \autoref{fig:sero_imm_c}).

\begin{sidewaysfigure}[H]
     \begin{center}
     \includegraphics[width=0.9\textwidth]{census_immunity_age.pdf}
     \end{center}
     \caption{The estimated immunity levels at the 2013 area unit level using the vaccine efficacy and serosurvey data presented in \autoref{table:sero}. See \autoref{fig:sero_imm_hist}, \autoref{fig:sero_imm_a}\autoref{fig:sero_imm_w}, \autoref{fig:sero_imm_c} for city maps and the distribution of immunity levels}
     \label{fig:sero_imm_map}
\end{sidewaysfigure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.9\textwidth]{census_immunity_age_Auk.pdf}
     \end{center}
     \caption{The estimated immunity levels at the 2013 area unit level for the Auckland vicinity using the vaccine efficacy and serosurvey data presented in \autoref{table:sero}.}
     \label{fig:sero_imm_a}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.9\textwidth]{census_immunity_age_Well.pdf}
     \end{center}
    \caption{The estimated immunity levels at the 2013 area unit level for the Wellington vicinity using the vaccine efficacy and serosurvey data presented in \autoref{table:sero}.}
     \label{fig:sero_imm_w}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.9\textwidth]{census_immunity_age_ChCh.pdf}
     \end{center}
    \caption{The estimated immunity levels at the 2013 area unit level for the Christchurch vicinity using the vaccine efficacy and serosurvey data presented in \autoref{table:sero}.}
     \label{fig:sero_imm_c}
\end{figure}

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.9\textwidth]{census_immunity_age_hist.pdf}
     \end{center}
     \caption{The distribution of estimated immunity levels at the 2013 area unit level using the vaccine efficacy and serosurvey data presented in \autoref{table:sero}. See \autoref{fig:sero_imm_map} for the plotted data}
     \label{fig:sero_imm_hist}
\end{figure}

@@

The regression analyses suggest that age is a particularly strong risk factor for measles. This comes as no surpise to epidemiologists or health care providers. However, our analyses also highlight other groups that are at greater risk of measles. In particular, Pacific people are at greater risk per capita, as are the less deprived (NZDep 1--3), and European and Maori 5--17 year old children compared to Asian and Pacific ethnicity children of the same age. Furthermore, Pacific origin people of all categories from 2--24 years old were less likely on a per capita basis to be cases than Europeans or Maori. Interpretation of these results must still be viewed with some caution, however, as there is very likely a spatial effect that might not be accounted for in these analyses. For example, increased numbers of more wealthy Europeans were measles cases, as was true for Maori. However, the alternative situation was clear for Pacific people, especially of the 0--1 year olds (\autoref{fig:percap_all} and \autoref{fig:cases_all}). Results such as this may simply be due to the spatial effects in the communities the outbreaks were occurring in, rather than an effect of wealth. The risk maps using the spatially explicit NIR vaccination data suggest that there is some heterogeneity in the vaccination cover at the area unit (a level above meshblock) for both MMR1 and MMR2 (see histograms within \autoref{fig:fig12006}, \autoref{fig:fig12007}, \autoref{fig:fig12008}, \autoref{fig:fig12009}, \autoref{fig:fig12010}, \autoref{fig:fig12011}, \autoref{fig:fig12012}, \autoref{fig:fig12013}, \autoref{fig:fig12014}, \autoref{fig:fig22006}, \autoref{fig:fig22007}, \autoref{fig:fig22008}, \autoref{fig:fig22009}, \autoref{fig:fig22010}, \autoref{fig:fig22011}, \autoref{fig:fig22012}, \autoref{fig:fig22013}, and \autoref{fig:fig22014}). This is also seen at the city level. The level of immunity within the population more broadly is also heterogeneous, however, it is noticable that the majority of area units have between 85 -- 96\% immunity, with few below this and few above this (\autoref{fig:sero_imm_hist}).

@@

\begin{figure}[H]
     \begin{center}
     \includegraphics[width=0.9\textwidth]{percap_eth_age.pdf}
     \end{center}
     \caption{Per capita measles case rates by prioritized ethnicity and age from 2007--2014. Full details are given in (\autoref{fig:percap_all})}
     \label{fig:PerCapAgeEth}
\end{figure}
%

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR2_NIR_Auk_2012.pdf}
 \end{center}
    \caption{Proportion of the population born in 2012 vaccinated against measles using MMR2 in each 2013 area unit in the Auckland vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig22012_a}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR2_NIR_Auk_2013.pdf}
 \end{center}
    \caption{Proportion of the population born in 2013 vaccinated against measles using MMR2 in each 2013 area unit in the Auckland vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig22013_a}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR1_NIR_Auk_2014.pdf}
 \end{center}
    \caption{Proportion of the population born in 2014 vaccinated against measles using MMR1 in each 2013 area unit in the Auckland vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig12014_a}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR2_NIR_Auk_2014.pdf}
 \end{center}
    \caption{Proportion of the population born in 2014 vaccinated against measles using MMR2 in each 2013 area unit in the Auckland vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig22014_a}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR2_NIR_Well_2012.pdf}
 \end{center}
    \caption{Proportion of the population born in 2012 vaccinated against measles using MMR2 in each 2013 area unit in the Wellington vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig22012_w}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR2_NIR_Well_2013.pdf}
 \end{center}
    \caption{Proportion of the population born in 2013 vaccinated against measles using MMR2 in each 2013 area unit in the Wellington vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig22013_w}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR1_NIR_Well_2014.pdf}
 \end{center}
    \caption{Proportion of the population born in 2014 vaccinated against measles using MMR1 in each 2013 area unit in the Wellington vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig12014_w}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR1_NIR_Well_2014.pdf}
 \end{center}
    \caption{Proportion of the population born in 2014 vaccinated against measles using MMR1 in each 2013 area unit in the Wellington vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig12014_w}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR2_NIR_Well_2014.pdf}
 \end{center}
    \caption{Proportion of the population born in 2014 vaccinated against measles using MMR2 in each 2013 area unit in the Wellington vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig22014_w}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR2_NIR_ChCh_2012.pdf}
 \end{center}
    \caption{Proportion of the population born in 2012 vaccinated against measles using MMR2 in each 2013 area unit in the Christchurch vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig22012_c}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR2_NIR_ChCh_2013.pdf}
 \end{center}
    \caption{Proportion of the population born in 2013 vaccinated against measles using MMR2 in each 2013 area unit in the Christchurch vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig22013_c}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR1_NIR_ChCh_2014.pdf}
 \end{center}
    \caption{Proportion of the population born in 2014 vaccinated against measles using MMR1 in each 2013 area unit in the Christchurch vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig12014_c}
\end{figure}

\begin{figure}
\begin{center}
    \includegraphics[width=0.9\textwidth]{nir_census_MMR2_NIR_ChCh_2014.pdf}
 \end{center}
    \caption{Proportion of the population born in 2014 vaccinated against measles using MMR2 in each 2013 area unit in the Christchurch vicinity. Note no adjustment is made for vaccine efficacy. See \autoref{table:tab_mmr_nir} and the supplementary data for details.}
% \end{center}
\label{fig:fig22014_c}
\end{figure}
